---
title: "mGPS - Soil microbiome results and figures"
output: 
  github_document
    
---

The the "mGPS.R" script and the "DATA" file containing Taxa abundacne data will both need to be in the working directory for this notebook these can be found at https://github.com/leomccarthy1/mGPS  


Dependencies
```{r, echo=T, eval = T}
library(sp)
library(rworldmap)
library(maps)
library(randomForest)
library(geosphere)
library(doParallel)
library(caret)
library(plyr)
library(maptools)
library(rgeos)
library(mapplots)
```

Load and combine, soil microbiome taxa and origin metadata

```{r}
environmental <- read.csv(file="DATA/Dataset_01_22_2018_enviro.csv",header=TRUE)
taxa <- read.csv(file="DATA/Dataset_01_22_2018_taxa.csv",header=TRUE)
SoilData <- merge(environmental,taxa,by.x="ID_Environmental",by.y="ID_Environmental")
```

Label country of each sample, some data cleaning. 
```{r}
SoilData$Country <- map.where(database="world", SoilData$Longitude, SoilData$Latitude)
SoilData[234:237,]$Country <- "Chile"
SoilData[c(2,6,9,15,22,67,34),]$Country <- "USA"

SoilData$Country <- factor(make.names(SoilData$Country))

#rename columns
colnames(SoilData)[3:5] <- c("longitude","latitude","continent")

#Remove countries with insufficient samples
keep <- levels(SoilData$Country)[table(SoilData$Country) > 2]
SoilData <- droplevels(SoilData[SoilData$Country%in% keep, ])

#concert read data to relative abundance.
for (i in 28:538){
  SoilData[,i] <- (SoilData[,i]/10000)
}

```


## Taxa selection

```{r}
#Remove redundant features
registerDoParallel(parallel::detectCores() - 1)
rfe_ctrl <- rfeControl(functions = rfFuncs,
                       method = "cv",
                       number = 5,
                       verbose = FALSE,
                       allowParallel = T
)
set.seed(12)
featureEliminationSoil <- rfe(x = SoilData[,c(28:538)],y = SoilData$Country,
                              sizes = c(20,30,50,100,200),
                              rfeControl = rfe_ctrl,
                              tuneLength = 5
)




#Extract and rank species importance

v <- varImp(featureEliminationSoil$fit, type = 1, scale = F)
v[,"taxa"] <- row.names(v)
v <- v[order(v$Overall,decreasing = T),]
top_species <- v[1:25,"taxa"]
```

Plot Fig S14, top 25 geographic indicator species

```{r}
par(font = 3)
dotchart(rev(v[1:25,"Overall"])*100,labels= rev(top_species),cex=1.2,pt.cex = 1.3,
         xlab="Mean decrease in accuracy", mgp = c(2.2,0,0))

```

load mGPS model
```{r}
source("mGPS.R")
```

Generate predictions through 5-fold cross validation
```{r}
#5 fold cross validation
set.seed(18)
trainFolds <-  createFolds(SoilData$Country, k = 5, returnTrain = T)

GeoPredsSoil <- list()
registerDoParallel(7) 

for (i in 1:5){
  
  train <- SoilData[trainFolds[[i]],]
  test <- SoilData[-trainFolds[[i]],]
  
  testPreds <-mGPS(training = train, testing = test, classTarget = "Country",variables = optimumVarsSoil)
  GeoPredsSoil[[i]] <- testPreds
  
}



add_preds <- list()
for (i in 1:5){
  
  add_preds[[i]] <- cbind(SoilData[-trainFolds[[i]],] , 
                          "countryPred"= GeoPredsSoil[[i]][[1]], 
                          "latPred" = GeoPredsSoil[[i]][[2]], 
                          "longPred" = GeoPredsSoil[[i]][[3]] )
  
  
  
}

SoilDataPreds <- rbind.fill(add_preds)

```

Pull predictions to nearest land mass

```{r}
#coast ajust#
#get world coastlines
coastlines <- cbind("x"  = SpatialLines2map(coastsCoarse)$x ,"y" =SpatialLines2map(coastsCoarse)$y)
coastlines <- coastlines[complete.cases(coastlines),]
coastlines <- coastlines[coastlines[,1] < 180 ,]

#Function to find the nearest land point of any GPS co-ordinates
find_coast <- function(long,lat){
  distances_from_coastline <-  spDistsN1(coastlines , c(long,lat), longlat = TRUE)
  
  closest_point <-  which.min(distances_from_coastline)
  new_coords <- coastlines[closest_point,]
  
  return(new_coords)
  
}


GPS_where <- map.where(database = "world", SoilDataPreds$longPred, SoilDataPreds$latPred)
toAdjust <- SoilDataPreds[which(is.na(GPS_where)),]
adjusted <- mapply(find_coast, long = toAdjust$longPred, lat = toAdjust$latPred )

SoilDataPreds[which(is.na(GPS_where)), "latPred"] <- adjusted[2,]
SoilDataPreds[which(is.na(GPS_where)), "longPred"] <- adjusted[1,]
```


Plot Fig 4, global soil microbiome predictions. 

```{r}
palette <-c( "deeppink2","darkorchid4","orangered2","gold2","brown","mediumspringgreen")
map <- getMap(resolution = "coarse")

plot(map,xlim = c(-160,160), col = "grey", border = "darkgrey", bg = "lightskyblue1", xlab = "", ylab = "")
title(ylab="Latitude",xlab = "Longitude", mgp=c(2,1,0),cex.lab=1.2)
for (i in 1:length(levels(SoilDataPreds$Continent))){{
  this_continent <- levels(SoilDataPreds$Continent)[i]
  find_lats <- SoilDataPreds[SoilDataPreds[,"Continent"] == this_continent,]$latPred
  find_longs <- SoilDataPreds[SoilDataPreds[,"Continent"] == this_continent,]$longPred
  points(find_longs, find_lats, col = palette[i], pch = "+", cex = 1.4)
  }
legend(-160,0,legend=c("Africa", "Asia", "Oceania", "Europe", "North America","South America"),
         col=palette,pch = "+",cex=1.4, bg = "lightskyblue1")  
}
map.axes()

```



Find the distance from country of origin for each sample. Put into data frame for plotting. 
```{r}
SoilDataPreds$GPSassign <- map.where(database = "world", SoilDataPreds$longPred, SoilDataPreds$latPred)
SoilDataPreds[is.na(SoilDataPreds$GPSassign),"GPSassign"] <- "In_sea"

for (i in 1:length((SoilDataPreds$Country))){ 
  country_lats <- map(database ="world",regions = SoilDataPreds[i,"Country"], plot = FALSE)$y
  country_longs <- map(database ="world",regions = SoilDataPreds[i,"Country"], plot = FALSE)$x
  pred_lat <- SoilDataPreds[i,]$latPred
  pred_long  <- SoilDataPreds[i,]$longPred
  distance <- c()
  if(SoilDataPreds[i,]$GPSassign == SoilDataPreds[i,]$Country){
    SoilDataPreds[i,"distance_from_country"] <- 0
  } else{
    for (n in 1:length(country_lats)){
      distance[n] <- distHaversine(c(pred_long ,pred_lat ),c(country_longs[n],country_lats[n]))/1000
      
    }
    SoilDataPreds[i,"distance_from_country"] <- min(distance, na.rm = TRUE)
  }
}

c <- confusionMatrix(SoilDataPreds$countryPred
                ,SoilDataPreds$Country)
c
mean(c[[4]][,"Sensitivity"])
R2(SoilDataPreds$longPred, SoilDataPreds$longitude)
mean(SoilDataPreds$distance_from_country < 100)

bar_df <- data.frame(row.names = c( "Overall",levels(SoilDataPreds$Country)))

for (i in 1: length(levels(SoilDataPreds$Country))){
  overall_prop <- mean(SoilDataPreds[,"distance_from_country"] < 500)
  bar_df[1,"0 - 500km"] <- overall_prop
  
  this_country <- levels(SoilDataPreds$Country)[i]
  prop <- mean(SoilDataPreds[SoilDataPreds$Country == this_country,][,"distance_from_country"] < 500)
  bar_df[i+1,"0 - 500km"] <- prop
}

for (i in 1: length(levels(SoilDataPreds$Country))){
  this_country <- levels(SoilDataPreds$Country)[i]
  prop <- mean(SoilDataPreds[SoilDataPreds$Country == this_country,][,"distance_from_country"] > 500 & SoilDataPreds[SoilDataPreds$Country == this_country,][,"distance_from_country"] < 1000)
  bar_df[i+1,"500 - 1000km"] <- prop
  
  overall_prop <- mean(SoilDataPreds[,"distance_from_country"] > 500 & SoilDataPreds[,"distance_from_country"] < 1000)
  bar_df[ 1,"500 - 1000km"] <- overall_prop
}

for (i in 1: length(levels(SoilDataPreds$Country))){
  this_country<- levels(SoilDataPreds$Country)[i]
  prop <- mean(SoilDataPreds[SoilDataPreds$Country == this_country,][,"distance_from_country"] > 1000 & SoilDataPreds[SoilDataPreds$Country == this_country,][,"distance_from_country"] < 2000)
  bar_df[i+1,"1000 - 2000km"] <- prop
  
  overall_prop <- mean(SoilDataPreds[,"distance_from_country"] > 1000 & SoilDataPreds[,"distance_from_country"] < 2000)
  bar_df[ 1,"1000 - 2000km"] <- overall_prop
}
for (i in 1: length(levels(SoilDataPreds$Country))){
  this_country <- levels(SoilDataPreds$Country)[i]
  prop <- mean(SoilDataPreds[SoilDataPreds$Country == this_country,][,"distance_from_country"] > 2000 & SoilDataPreds[SoilDataPreds$Country == this_country,][,"distance_from_country"] < 3000)
  bar_df[i+1,"2000 - 3000km"] <- prop
  
  overall_prop <- mean(SoilDataPreds[,"distance_from_country"] > 2000 & SoilDataPreds[,"distance_from_country"] < 3000)
  bar_df[ 1,"2000 - 3000km"] <- overall_prop
}
for (i in 1: length(levels(SoilDataPreds$Country))){
  this_country <- levels(SoilDataPreds$Country)[i]
  prop <- mean(SoilDataPreds[SoilDataPreds$Country == this_country,][,"distance_from_country"] > 3000 )
  bar_df[i+1,"> 3000km"] <- prop
  
  overall_prop <- mean(SoilDataPreds[,"distance_from_country"] > 3000)
  bar_df[ 1,"> 3000km"] <- overall_prop
}
size1 <- c()
for (i in 1: length(levels(SoilDataPreds$Country))){
  
  this_country <- levels(SoilDataPreds$Country)[i]
  size1[i] <- length(which(SoilDataPreds$Country == this_country))
}


```


Plot Fig S8, distance from country of origin barplot

```{r}
par(xpd = T, mar = par()$mar + c(1,0,0,7), mgp = c(0,0.7,0), las=2)
bp <- barplot(t(bar_df*100), col=c("slategray1","lightblue", "skyblue", "royalblue3", "darkblue"), 
              names.arg=c("Overall",paste0(levels(SoilDataPreds$Country),"  (",size1,")")) ,
              args.legend = list(x = "topright", inset=c(-0.5,0)), las =2, 
              cex.names=.6,ylab = "", axisnames = F,axes = F, space =0)

axis(side =2, pos = 0)
mtext(text = c("Overall",paste0(levels(SoilDataPreds$Country),"  (",size1,")")) , side = 1, at = bp, line = 0, padj = 1, cex = 0.7)
title(ylab="Proportion of sample predictions %", mgp=c(0,0,0),cex.lab=1)
legend("topright",inset = c(-0.1,0.4), rev(c(colnames(bar_df))), fill = rev(c("slategray1","lightblue", "skyblue", "royalblue3", "darkblue")) , bty = 1, cex = 0.8)

par(mar=c(5, 4, 4, 2) + 0.1)

```



Generate Fig S15, plots of abundance geo distribution for 25 geographic indicator species. These plots will be saved to the working directory. 

```{r}
ag <- aggregate(SoilDataPreds[,top_species], by = list(SoilDataPreds$Country), FUN = median)
country_locations <- aggregate(SoilDataPreds[,c("longitude","latitude")], by = list(SoilDataPreds$Country), FUN = mean)
country_abund <- merge(country_locations,ag, by = "Group.1")

for (i in top_species){
  if (max(country_abund[,i]) == 0 ){country_abund[,i] <- 0} 
  else{ country_abund[,i] <- (country_abund[,i]- min(country_abund[,i]) )/(max(country_abund[,i])- min(country_abund[,i]))
  }
}

pal <- palette <-c( "gold2","brown","dodgerblue3","darkseagreen2","darkorchid4","darkcyan","orangered2","olivedrab2","deeppink2","darkslateblue","mediumspringgreen","gray21","royalblue","yellow1","orange","purple1","cyan2","burlywood","aquamarine2","chartreuse4","deeppink4","cadetblue4","goldenrod1","firebrick2","hotpink")
for (p in 1:25){
  
  
  #png(paste("Species_abundance_",p,".png", sep = ""), width = 12, height = 8, units = "in", res = 600)
  
  ggplot(map, aes(x = long, y = lat, group = group)) +
    geom_polygon(fill="grey", colour = "darkgrey", size = 0.4)+
    coord_quickmap(xlim = c(-145,165), ylim =c(-50,70))+
    theme_classic()+
    theme(panel.background = element_rect(fill = "lightskyblue1",
                                          colour = "lightskyblue1",
                                          size = 0.5, linetype = "solid"))+
    geom_point(data = country_abund , aes(x = longitude,y = latitude, 
                                      group = Group.1, color = country_abund[,p+3]), size = 9)+
    
    scale_color_gradient(low = "white", high = pal[p], na.value = NA,name = "Relative abundance \n(normalised)", limits = c(0,1), breaks = waiver(),n.breaks = 6)+
    
    
    theme(legend.position = c(0.08, 0.3),
          legend.background = element_rect(fill="lightskyblue1", 
                                           size=2, linetype="solid"),legend.key.size = unit(0.9, "cm"), 
          legend.text = element_text(size = 9, face = "bold"), 
          legend.title = element_text(size = 12, face = "bold"),
          plot.title = element_text(size = 18, face = "bold"),
          axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          plot.margin = unit( c(t = 0, r = -.1, b = -.2, l = 0), "cm"))+
    ggtitle(colnames(country_abund )[p+3],)+
    labs(x="", y="")
  
  ggsave(paste("Species_abundance_soil_",p,".png", sep = ""),plot = last_plot(),device = NULL,path = NULL,scale = 1,width = 12,height = 5.1,dpi = 600,limitsize = TRUE,)
  
  
}

```


